---
title: "Get and filter open data using phsopendata getresource()"
author: "Pauline Ward"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(dplyr)
library(phsopendata)
library(openxlsx)
library(here)

```

# Getting data using phsopendata 

Cancer mortality data is available from opendata.nhs.scot. 

Data is filtered by year and site during the download step. 

Then I've done a join on the local authority (East Renfrewshire in the first 
instance), so that I can select LA 
transparently by name, rather than hardwiring in a code. 

```{r get_council_areas, echo=TRUE}
# Geographic data - local authority council areas
# https://www.opendata.nhs.scot/dataset/geography-codes-and-labels/resource/967937c4-8d67-4f39-974f-fd58c4acfda5
Councils_resource_id <- "967937c4-8d67-4f39-974f-fd58c4acfda5"
tbl_Council_Areas <- get_resource(
  Councils_resource_id, 
  col_select = c("_id", "CA", "CAName", "CADateArchived")
)

tbl_Council_Areas <- tbl_Council_Areas |>
  filter(is.na(CADateArchived)) |>
  unique()

write.xlsx(tbl_Council_Areas, "Council_names_and_codes.xlsx") 

```
## Local authority council area names
I have all the councils ```{r tbl_Council_areas$CAName}``` right here. 

```{r get_and_filter_data, echo=TRUE}
councils_of_interest <- c("East Renfrewshire", "Aberdeenshire")

## Address for importing mortality data 
## ... from https://www.opendata.nhs.scot/dataset/cancer-mortality/resource/eebc8f38-7297-4bdc-a417-69ce3e2e6d44

# ID of dataset: mortality by council area
resource_id <- "eebc8f38-7297-4bdc-a417-69ce3e2e6d44" 
# this_data <- get_resource_sql()
this_data <- get_resource(
  resource_id, 
                          # comment out col_select, to just get all the columns
                          # col_select = c("CA", "CancerSiteICD10Code", "CancerSite", "Sex", "Year", "DeathsAllAges")
  row_filters = list(
                    Year = "2020", 
                    CancerSite = "Breast"
  )
)

this_data <- this_data |>
  filter(Sex == "All")

# this_data is a tibble
class(this_data)

# Add Council Area names to the data
this_data <- this_data |>
  left_join(tbl_Council_Areas, by = "CA")

# Filter to the council area(s) of interest
tbl_relevant_data <- this_data |>
  filter(CAName %in% councils_of_interest)

```

# Results
